---
title: "Index.qmd"
author: "Faith Rhinehart"
format: 
  html:
    embed-resources: true
---
##Packages and reading in the datasets
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
pixar_films <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv')
public_response <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv')

pixarMovies <- merge(pixar_films, public_response, by = "film") |> select(1, 3:9) |> slice(1:10, 12:24)
```

## Finding highest score
```{r}
#Highest rotten tomatoes score
high_rotten <- pixarMovies |> arrange(desc(rotten_tomatoes)) |> select(1,5) |>
  mutate(Movie = fct_reorder(film, rotten_tomatoes))
```

```{r}
#Highest meta critic score
high_meta <- pixarMovies |> arrange(desc(metacritic)) |> select(1,6) |>
  mutate(Movie = fct_reorder(film, metacritic))
View(high_meta)
```

```{r}
#Highest critics choice score
high_critic <- pixarMovies |> arrange(desc(critics_choice)) |> select(1,8) |> slice(1:21) |>
  mutate(Movie = fct_reorder(film, critics_choice))
View(high_critic)
```

##Visualizations
```{r}
#Rotten Tomatoes
ggplot(data = high_rotten, aes(x = Movie, y = rotten_tomatoes, fill = rotten_tomatoes)) +
  geom_col() +
  coord_flip() +
  labs(y = "Rotten Tomatoes Score",
       title = "Rotten Tomatoes Score for Pixar Movies") +
  scale_fill_viridis_c() +
  theme_minimal()
```

```{r}
#Meta Critic
ggplot(data = high_meta, aes(x = Movie, y = metacritic, fill = metacritic)) +
  geom_col() +
  coord_flip() +
  labs(y = "Metacritic Score",
       title = "Metacritic Score for Pixar Movies") +
  scale_fill_viridis_c() +
  theme_minimal()
```

```{r}
#Critics Choice
ggplot(data = high_critic, aes(x = Movie, y = critics_choice, fill = critics_choice)) +
  geom_col() +
  coord_flip() +
  labs(
    y = "Critic's Choice Score",
    title = "Critic's Choice Score for Pixar Movies") +
  scale_fill_viridis_c() +
  theme_minimal()
```

Introduction:
what data set are you working with?

_Pixar Films descriptions and public response to each film. Two data sets were combined into one_

what variables are you most interested in? How many observations are in your data set?

_Variables interested in: film, rotten tomatoes, metacritic, and critics choice_

_There are 24 observations with 8 variables_

what is your question of interest?

_Which films has the overall highest score for rotten tomatoes, meta critics and critics choice_

where did you get the data from? (If possible, give a link to the data source).

_Link to both data sets: Link to data set: https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-03-11/readme.md_

what other pieces of context are important to explain to the reader?

_That a few movies don't have a critics choice score so when that visual is made those observations will be taken out of the data set and noted. Luca doesn't have a score for any of the three so it will be taken out of the dataset._  

Primary Visualizations
choose an appropriate visualization, eliminate clutter, and focus attention as appropriate to answer your question of interest.

provide a written interpretation of each visualization you have, explaining what a reader should “take away” from each visualization.

Conclusion and Wrap-Up
are there any flaws with your approach to analyzing the data you’ve selected?

are there future directions you would go if you had additional data or more time?

Connection to Class Ideas
At the end of your write-up, you should connect your visualization(s) to ideas from class. In what ways would you argue your visualization(s) are effective ways of communicating information?